---
title: "Column Modification - \"Country\""
author: "Charles J. Gomez, Harshvardhan Singh"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The country column in our dataset contains a more detailed information about the authors' affiliations, including first, middle, and last authors. However, this brings a level of complexity and potential confusion for machine learning models. Therefore, we propose a method to simplify the country column by retaining only the first and last author affiliations, ensuring a more straightforward representation of authorship.
```{r}
library(stringr)

data = read.csv("INPUT_SQL_Text_Data_Astronomy_and_Astrophysics.csv")
data_CountryColumn <- data$country

#We removed all instances of middle authors' country codes. This step involved replacing the pattern "\w+\+middle" with an empty string, effectively eliminating all middle author affiliations. The resulting country codes were then trimmed to remove any leading or trailing spaces

data$country <- gsub("\\w+\\+middle", "", data$country)
data$country <- trimws(data$country)
print(head(data$country, 50))

#We removed the "+first" and "+last" strings from the remaining country codes using another regular expression substitution. The pattern "\+last|\+first" was replaced with an empty string, ensuring that only the country codes remained

data$country <- gsub("\\+last|\\+first", "", data$country)
# Remove any extra spaces resulting from the removal
data$country <- sapply(strsplit(data$country, "\\s+"), function(x) paste(x[x != ""], collapse = " "))

print(head(data$country, 50))



# Function to calculate the percentage count of each country code in a vector.
calculate_percentage <- function(vec) {
  counts <- table(vec)
  percentages <- prop.table(counts) * 100
  formatted <- paste0(round(percentages, 1), "%", names(percentages))
  paste(formatted, collapse = " ")
}

# Apply the calculate_percentage function to each row in the 'country' column
data$country <- sapply(strsplit(data$country, "\\s+"), calculate_percentage)

# Print the modified 'country' column (first 20)
print(head(data$country, 50))

#print first 50 rows of the dataset
#print(head(data, 50))
```
