---
title: "Preprocessing"
author: "Charles J. Gomez, Harshvardhan Singh"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The country column in our dataset contains a more detailed information about the authors' affiliations, including first, middle, and last authors. However, this brings a level of complexity and potential confusion for machine learning models. Therefore, we propose a method to simplify the country column by retaining only the first and last author affiliations, ensuring a more straightforward representation of authorship.

Moreover, we collapse the countries (except the top 15 countries) into the following regions -
North/South America
Europe
Asia/Oceania
Africa

List of Countries-

**Top 15 - 91.47% of global contribution (New Data)** 
1. USA (US) - 
2. China (CN) - 
3. Great Britain (GB) - 
4. Germany (DE) -
5. Japan (JP) -
6. France (FR)
7. Italy (IT)
8. 

**Top 15 - 91.47% of global contribution (Old Data)** 
1. USA (US) - 50.61%
2. Great Britain (GB) - 10.13%
3. Germany (DE) - 5.73%
4. Italy (IT) - 3.84%
5. Australia (AU) - 2.82%
6. Canada (CA) - 2.67%
7. Japan (JP) - 2.59%
8. China (CN) - 2.18%
9. Netherlands (NL) - 2.01%
10. France (FR) - 1.98% 
11. Russia (RU) - 1.85% 
12. Spain (ES) - 1.78% 
13. India (IN) - 1.42% 
14. Israel (IL) - 0.96% 
15. Switzerland (CH) - 0.90%


**The Americas (North/South America)**
1. BR - Brazil
2. CL - Chile
3. VE - Venezuela
4. MX - Mexico
5. CO - Colombia
6. AR - Argentina
7. UY - Uruguay
8. JM - Jamaica
9. CR - Costa Rica
10. GT - Guatemala
11. HN - Honduras
12. PR - Puerto Rico
13. VI - U.S. Virgin Islands
14. BO - Bolivia
15. PE - Peru
16. EC - Ecuador
17. CU - Cuba

**Europe**
1. RS - Serbia
2. DK - Denmark
3. SE - Sweden
4. BE - Belgium
5. PL - Poland
6. IE - Ireland
7. FI - Finland
8. AT - Austria
9. HU - Hungary
10. LV - Latvia
11. UA - Ukraine
12. GE - Georgia
13. GR - Greece
14. HR - Croatia
15. TR - Turkey
16. AM - Armenia
17. CZ - Czech Republic
18. PT - Portugal
19. SK - Slovakia
20. IS - Iceland
21. NO - Norway
22. BG - Bulgaria
23. SI - Slovenia
24. EE - Estonia
25. CY - Cyprus
26. LU - Luxembourg
27. MT - Malta
28. MK - North Macedonia
29. RO - Romania
30. LT - Lithuania

**Africa**
1. NG - Nigeria
2. MZ - Mozambique
3. BW - Botswana
4. ZA - South Africa
5. GH - Ghana
6. RW - Rwanda
7. UG - Uganda
8. EG - Egypt
9. TZ - Tanzania
10. MU - Mauritius
11. ZW - Zimbabwe
12. NA - Namibia
13. SD - Sudan
14. TN - Tunisia
15. BD - Burundi

**Asia/Oceania**
1. KR - South Korea
2. TW - Taiwan
3. IR - Iran
4. KZ - Kazakhstan
5. KH - Cambodia
6. PK - Pakistan
7. VN - Vietnam
8. MY - Malaysia
9. SA - Saudi Arabia
10. UZ - Uzbekistan
11. TH - Thailand
12. NP - Nepal
13. LK - Sri Lanka
14. ID - Indonesia
15. TJ - Tajikistan
16. AE - United Arab Emirates
17. LB - Lebanon
18. SG - Singapore
19. AZ - Azerbaijan
20. IQ - Iraq
21. OM - Oman
22. KW - Kuwait
23. QA - Qatar
24. NZ - New Zealand
25. PS - Palestine
26. BD - Bangladesh

```{r}
library(stringr)

data = read.csv("filtered_data_feb_2024.csv")
data <- subset(data, year >= 1990 & year <= 2015)
data_CountryColumn <- data$country

data$concatenated_title_abstract <- paste(data$title, data$paperabstract, sep = " ")

#We removed all instances of middle authors' country codes. This step involved replacing the pattern "\w+\+middle" with an empty string, effectively eliminating all middle author affiliations. The resulting country codes were then trimmed to remove any leading or trailing spaces

data$country <- gsub("\\w+\\+middle", "", data$country)
data$country <- trimws(data$country)
print(head(data$country, 50))

#We removed the "+first" and "+last" strings from the remaining country codes using another regular expression substitution. The pattern "\+last|\+first" was replaced with an empty string, ensuring that only the country codes remained

data$country <- gsub("\\+last|\\+first", "", data$country)
# Remove any extra spaces resulting from the removal
data$country <- sapply(strsplit(data$country, "\\s+"), function(x) paste(x[x != ""], collapse = " "))

print(head(data$country, 50))



# calculate the percentage count of each country code in a vector
#calculate_percentage <- function(vec) {
#  counts <- table(vec)
#  percentages <- prop.table(counts) * 100
#  formatted <- paste0(round(percentages, 1), "%", names(percentages))
#  paste(formatted, collapse = " ")
#}

# Apply the calculate_percentage function to each row in the 'country' column
#data$country <- sapply(strsplit(data$country, "\\s+"), calculate_percentage)

# Print the modified 'country' column (first 20)
#print(head(data$country, 50))




# Calculate percentage count of each country code in a vector
calculate_percentage <- function(vec) {
  counts <- table(vec)
  percentages <- prop.table(counts) * 100
  formatted_percentages <- round(percentages, 2)
  return(percentages)
}

# Apply the calculate_percentage function to each row in the 'country' column
percentage_counts <- lapply(strsplit(data$country, "\\s+"), calculate_percentage)

# Get all unique countries
all_countries <- unique(unlist(lapply(percentage_counts, names)))

# Initialize a new data frame to hold the percentages
percentage_df <- data.frame(matrix(ncol = length(all_countries), nrow = length(percentage_counts)))
names(percentage_df) <- all_countries

# Fill the data frame with percentages
for (i in seq_along(percentage_counts)) {
  country_names <- names(percentage_counts[[i]])
  country_percentages <- percentage_counts[[i]]
  percentage_df[i, country_names] <- country_percentages
}

# Replace NA values with 0
percentage_df[is.na(percentage_df)] <- 0

# Combine the original data with the new percentage-encoded country columns
data <- cbind(data, percentage_df)

# Summary of dataframe
summary(data)

# structure of the dataframe
str(data)

```

```{r}

# Load the necessary libraries
library(ggplot2)
library(dplyr)

#Histogram by number of articles each country has contributed to

# Calculate count of each country code in a vector
calculate_count <- function(vec) {
  counts <- table(vec)
  return(counts)
}

# Apply the calculate_count function to each row in the 'country' column
counts <- lapply(strsplit(data$country, "\\s+"), calculate_count)

# Get all unique countries
all_countries <- unique(unlist(lapply(counts, names)))

# Initialize a new data frame to hold the counts
count_df <- data.frame(matrix(ncol = length(all_countries), nrow = length(counts)))
names(count_df) <- all_countries

# Fill the data frame with counts
for (i in seq_along(counts)) {
  country_names <- names(counts[[i]])
  country_counts <- counts[[i]]
  count_df[i, country_names] <- country_counts
}

# Replace NA values with 0
count_df[is.na(count_df)] <- 0

# Combine all countries into one column for the histogram
all_countries_df <- stack(count_df)

# Rename the columns
colnames(all_countries_df) <- c("Count", "Country")

# Remove rows where count is zero
all_countries_df <- all_countries_df[all_countries_df$Count > 0,]

# Calculate the total counts for each country
all_countries_df <- all_countries_df %>%
  group_by(Country) %>%
  summarise(Total = sum(Count)) %>%
  arrange(desc(Total))

# Split the data frame into 5 equal parts
split_data <- split(all_countries_df, cut(seq(nrow(all_countries_df)), 5, labels = FALSE))

# Create a list to store the plots
plot_list <- list()

# Iterate over each subset of data and create a histogram with count
plot_list <- lapply(1:5, function(i) {
  ggplot(split_data[[i]], aes(x=reorder(Country, -Total), y=Total)) +
    geom_bar(stat="identity", fill="steelblue") +
    geom_text(aes(label=Total), vjust=-0.5, size=2.5) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x="Country", y="Number of Contributions", 
         title = paste("Research Publications by Country: Part", i))
})

```

```{r}

# View each plot by calling it from the list
plot_list[[1]]
plot_list[[2]]
plot_list[[3]]
plot_list[[4]]
plot_list[[5]]
```

```{r}

#Histogram of total research contributions by country


# Combine all countries into one column for the histogram
all_countries_df <- stack(percentage_df)

# Rename the columns
colnames(all_countries_df) <- c("Percentage", "Country")

# Remove rows where percentage is zero
all_countries_df <- all_countries_df[all_countries_df$Percentage > 0,]

# Calculate the total percentage for each country
all_countries_df <- all_countries_df %>%
  group_by(Country) %>%
  summarise(Total = sum(Percentage)) %>%
  arrange(desc(Total))

# Split the data frame into 5 equal parts
split_data <- split(all_countries_df, cut(seq(nrow(all_countries_df)), 5, labels = FALSE))

# Create a list to store the plots
plot_list <- list()

# Define the text size for each plot
text_size <- c(1.5, 2, 2, 2.5, 2.5)

# Iterate over each subset of data and create a histogram
plot_list <- lapply(1:5, function(i) {
  ggplot(split_data[[i]], aes(x=reorder(Country, -Total), y=Total)) +
    geom_bar(stat="identity", fill="purple") +
    geom_text(aes(label=sprintf("%.2f", Total)), vjust=-0.5, size=text_size[i]) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x="Country", y="Total Contribution Count", 
         title = paste("Total Research Contributions (total weight): Part", i))
})

```

```{r}

# View each plot by calling it from the list
plot_list[[1]]
plot_list[[2]]
plot_list[[3]]
plot_list[[4]]
plot_list[[5]]

```

```{r}

# Histogram of Research Contributions % by Country

# Calculate total of all percentages
total_percentage <- sum(all_countries_df$Total)

# Express each country's contribution as a percentage of total
all_countries_df <- all_countries_df %>%
  mutate(Percentage_of_Total = Total / total_percentage * 100) %>%
  arrange(desc(Percentage_of_Total))

# Split the data frame into 5 equal parts
split_data <- split(all_countries_df, cut(seq(nrow(all_countries_df)), 5, labels = FALSE))

# Create a list to store the plots
plot_list <- list()

# Define the decimal places for each plot
formats <- c("%.2f%%", "%.2f%%", "%.3f%%", "%.4f%%", "%.4f%%")

# Define the text size for each plot
text_size <- c(2.5, 2.5, 2, 2, 2)

# Iterate over each subset of data and create a histogram
plot_list <- lapply(1:5, function(i) {
  ggplot(split_data[[i]], aes(x=reorder(Country, -Percentage_of_Total), y=Percentage_of_Total)) +
    geom_bar(stat="identity", fill="orange") +
    geom_text(aes(label=sprintf(formats[i], Percentage_of_Total)), vjust=-0.5, size=text_size[i]) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(x="Country", y="Total Contribution", 
         title = paste("Percentage Distribution of Global Research Contributions: Part", i))
})

```

```{r}

# View each plot by calling it from the list
plot_list[[1]]
plot_list[[2]]
plot_list[[3]]
plot_list[[4]]
plot_list[[5]]

```

```{r}

# List the country codes for each region

north_america <- c('MX', 'PR', 'CU', 'DO', 'GT', 'HN', 'CR', 'PA', 'NI', 'SV', 'BZ', 'JM', 'TT', 'BS', 'VG', 'GD', 'BB', 'KN', 'GP', 'GF', 'MS')
south_america <- c('BR', 'AR', 'CL', 'CO', 'VE', 'PE', 'EC', 'UY', 'BO')
europe <- c('RO', 'BE', 'GR', 'CZ', 'PT', 'SE', 'HU', 'BY', 'AT', 'CH', 'DK', 'FI', 'SK', 'NO', 'IE', 'HR', 'SI', 'LV', 'LT', 'EE', 'IS', 'MT', 'LU', 'MC', 'VA', 'RS', 'ME', 'MK', 'AL', 'BA', 'XK', 'UA', 'MD', 'BG', 'CY', 'TR', 'GE', 'AM', 'AZ', 'LI', 'SJ', 'JE', 'BM', 'SX', 'GL')
africa <- c('DZ', 'AO', 'EG', 'LY', 'MA', 'SD', 'TN', 'ET', 'KE', 'TZ', 'UG', 'DZ', 'NG', 'GH', 'CI', 'SN', 'CD', 'ZW', 'RW', 'ZM', 'MG', 'MW', 'AO', 'CM', 'BF', 'NE', 'MU', 'MZ', 'LS', 'BW', 'GM', 'TD', 'CG', 'NA', 'SS', 'SZ', 'BJ', 'TG', 'CI', 'ST', 'BI', 'RE', 'ZA')
asia <- c('ID', 'TH', 'PK', 'MY', 'PH', 'VN', 'LK', 'BD', 'NP', 'MM', 'TW', 'HK', 'SG', 'MN', 'KH', 'BT', 'MO', 'BN', 'MV', 'KW', 'AF', 'AE', 'SY', 'TJ', 'KZ', 'KG', 'IL', 'PS', 'LB', 'KP', 'BH', 'IQ', 'YE', 'JO', 'IR', 'QA', 'OM', 'UZ', 'SA')
oceania <- c('NZ', 'PG', 'FJ', 'VU', 'NC', 'PF', 'WS', 'TO')


# Create the regional sum columns
data$NorthAmerica <- rowSums(data[,north_america], na.rm = TRUE)
data$SouthAmerica <- rowSums(data[,south_america], na.rm = TRUE)
data$Europe <- rowSums(data[,europe], na.rm = TRUE)
data$Africa <- rowSums(data[,africa], na.rm = TRUE)
data$Asia <- rowSums(data[,asia], na.rm = TRUE)
data$Oceania <- rowSums(data[,oceania], na.rm = TRUE)

# drop the original country columns
country_cols <- c(north_america, south_america, europe, africa, asia, oceania)
data <- data[, !(names(data) %in% country_cols)]



# Select the columns of interest
selected_countries <- c('US', 'GB','DE', 'CN','JP', 'FR', 'IT','CA','ES', 'RU', 'AU','NL', 'IN','KR', 'PL', 'NorthAmerica', 'SouthAmerica', 'Europe', 'Africa', 'Asia', 'Oceania')
selected_data <- data[selected_countries]

# Histogram by number of contributions each country or region has contributed to
all_countries_df <- stack(selected_data)
colnames(all_countries_df) <- c("Count", "Country")
all_countries_df <- all_countries_df[all_countries_df$Count > 0,]

# Calculate the total counts for each country
all_countries_df <- all_countries_df %>%
  group_by(Country) %>%
  summarise(Total = sum(Count)) %>%
  arrange(desc(Total))

# Histogram of total research contributions by country
all_countries_df <- stack(selected_data)
colnames(all_countries_df) <- c("Count", "Country")
all_countries_df <- all_countries_df[all_countries_df$Count > 0,]

all_countries_df <- all_countries_df %>%
  group_by(Country) %>%
  summarise(Total = sum(Count)) %>%
  arrange(desc(Total))

# Histogram of Research Contributions % by Country
total_percentage <- sum(all_countries_df$Total)
all_countries_df <- all_countries_df %>%
  mutate(Percentage_of_Total = Total / total_percentage * 100) %>%
  arrange(desc(Percentage_of_Total))

# Create the plot
ggplot(all_countries_df, aes(x=reorder(Country, -Percentage_of_Total), y=Percentage_of_Total)) +
  geom_bar(stat="identity", fill="red") +
  geom_text(aes(label=sprintf("%.2f%%", Percentage_of_Total)), vjust=-0.5, size=2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(x="Country/Region", y="Percentage of Total Contributions", 
       title = "Percentage Distribution of Global Research Contributions")


# Find the smallest (earliest) year
earliest_year <- min(data$year, na.rm = TRUE)
print(paste("The earliest publication year is:", earliest_year))

# Find the greatest (latest) year
latest_year <- max(data$year, na.rm = TRUE)
print(paste("The latest publication year is:", latest_year))

#Create table for the count of publications in each year
table(data$year)

#Plot full table
ggplot(data, aes(x = year)) +
  geom_bar(fill = "steelblue") +
  theme_minimal() +
  labs(x = "Year", y = "Number of papers", title = "Number of papers per year")




write.csv(data, "preprocessed_data_Feb_2024.csv", row.names = FALSE)

```
